---
title: "Data.table"
output: github_document
date: "<em style='color:#00508060;'>`r Sys.setlocale('LC_ALL','English'); format(Sys.time(), '%B %d, %Y')`</em>"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, results='hold', collapse=TRUE)
options(datatable.print.nrows=6, datatable.print.trunc.cols=TRUE)
```

- Example data

```{r}
library(data.table)

dt_starwars <- data.table(dplyr::starwars)
dt_starwars
```

- `DT[i, j, by]` For SQL users: `from[where, select, group by]`
- Vectors & data types, lists, data frames, data tables

## Actions on rows: `DT[i]`

#### Sorting with `DT[order(...)]`

A data.table that has been sorted on one or more variables, in ascending or descending order, can be returned by using the `order()` function. The `order()` function returns the numerical index value of elements after sorting the data. For example, with the vector `abc <- c('a', 'c', 'b')` we get `1 3 2` when we execute `order(abc)`.

<details>
  <summary>
    Return a data.table sorted by a single variable in <em>ascending</em> order:
  </summary>
  ```{r}
  dt_starwars[order(height)]
  ```
</details>

<details>
  <summary>
    Return a data.table sorted by a single variable in <em>descending</em> order:
  </summary>
  ```{r}
  dt_starwars[order(-height)]
  ```
</details>

<details>
  <summary>
    Return a data.table sorted by <em>multiple</em> variables:
  </summary>
  ```{r}
  dt_starwars[order(sex, height)]
  ```
</details>

-----

### To add:

- filter (numeric); extend on sort to get top n
- filter (logical); logical operators
- select; `dt[, j]`, `.SD`
- assignment
- conditional assignment (`dt[, ifelse(...)]` and `dt[i, j]`)
- compute; `x(j)`
- group by; `by` and `keyby`
- special characters; `.SD, .N, .I ...`
- chaining

<!--


#### Subset

###### Logical/Boolean

```{r}
dt_starwars[height>200]
dt_starwars[sex=='Male' & height>200]
dt_starwars[height>=200 & height<=300]
```

###### Positional

```{r}
dt_starwars[1]
dt_starwars[1:5]
dt_starwars[c(1, 10)]
```

-->
